---
title: Estimators
author: George
date: '2020-08-27'
slug: estimators
categories: []
tags: []
---

```{r echo = F, warning = F}
library(data.table)
library(ggplot2)
library(animation)
library(gifski)
theme_set(theme_light())
```

## _Why do we need an estimator?_

In the previous [post](/2020/08/25/first-post/) we discussed statistical models and the idea that a statistical model is a pair ${\left(S, \{ P_\theta \} _{\theta \in \Theta }\right)}$. We also discussed the fact that if our model is well specified then the true distribution (the one that governs reality) will be included in our selected family of distributions $P_{\theta \in \Theta}$ for certain $\theta$. But we did not discuss at all how we can estimate this $\theta$. This is were *estimators* come into play.    

Suppose we have a true distribution with the true parameter $\color{blue}\theta$ and we can obtain data from that distribution. Our goal is to be able to "guess" what that parameter is by using the data. In particular, what we need is a function that takes as input data from the true distribution and outputs a "guess" or an estimate for the true parameter $\color{blue}\theta$.

fd

```{r, animation.hook="gifski"}
for (i in 1:5) {
  hist(rpois(i*2000, lambda = 2), freq = T, breaks = 0:10)
  #pie(c(i %% 2, 6), col = c('red', 'yellow'), labels = NA)
}
```




