---
title: Estimators
author: George
date: '2020-08-27'
slug: estimators
categories: []
tags: []
---

```{r echo = F, warning = F}
library(data.table)
library(ggplot2)
library(animation)
library(gifski)
theme_set(theme_light())
```

## _Why do we need an estimator?_

In the previous [post](/2020/08/25/first-post/) we discussed statistical models and the idea that a statistical model is a pair ${\left(S, \{ P_\theta \} _{\theta \in \Theta }\right)}$. We also discussed the fact that if our model is well specified then the true distribution (the one that governs reality) will be included in our selected family of distributions $P_{\theta \in \Theta}$ for certain $\theta$. But we did not discuss at all how we can estimate this $\theta$. This is were *estimators* come into play.    

Suppose we have a true distribution with the true parameter $\color{blue}\theta$ and we can obtain sample data from that distribution. Our goal is to be able to estimate quantities of the true distribution such as $\color{blue}\theta$, its mean or its median  by using the sample data. In particular, what we need is a function that takes as input the sample data from the true distribution and outputs a "guess" or an estimate for the quantity of the true distribution we are interested in.

As an example, suppose the true distribution is a Poisson with lambda = 2 (**lambda is unknown to us**) and we can obtain sample data as seen in the picture below.  
_What would be a good guess for lambda given the sample data?_

```{r, animation.hook="gifski", echo = F}
y_max = 2000
full_data = rpois(4000, lambda = 2)
for (i in 1:10) {
  data = full_data[1:(i*400)]
  temp = hist(data, freq = T, breaks = 0:12, xaxt="n", ylim=c(0,y_max), main = paste0(i*400, " samples from a Poisson with lambda = 2"), xlab = "")
  text(x = 10, y = y_max, labels = "Sample statistics:")
  text(x = 10, y = y_max-150, labels = paste0("Sample mean = ", round(mean(data), 2)))
  text(x = 10, y = y_max - 300, labels = paste0("Sample variance = ", round(var(data), 2)))
  text(x = 10, y = y_max - 450, labels = paste0("Sample max = ", round(max(data), 2)))
  text(x = 10, y = y_max - 600, labels = paste0("Sample min = ", round(min(data), 2)))
  axis(side=1,at=temp$mids,labels= 0:11)
}
```
To answer this question we have to think a bit about our distribution. If $X$ follows a Poisson distribution with parameter lambda, then we know that the mean of $X$ equals lambda ($E[X] = \lambda$). That gives a hint as to what would be a good estimator for lambda, the sample mean. And looking at the chart above, we see that as the sample size goes bigger, the sample mean tends goes to 2 which is indeed the lambda of the true distribution.  

But are we sure that it is a good estimator? What if it was by chance? Can we quantify how good an estimator is?




